\subsection{Speculative Poisoned FIFO}\label{sec:specpoisonfifo}

Entries in the speculative poisoned FIFO carry speculation bit masks, but they are marked as \emph{poisoned} instead of being dropped when mis-speculation happens.
This is commonly used skid FIFOs in parallel with in-order functional pipelines that we don't know details, e.g., the floating point pipeline.

\subsubsection{Interface}
Figure~\ref{fig:specpoisonfifo-ifc} shows the interface of the speculative poisoned FIFO (type \code{ToSpecFifo} has been explained in Figure~\ref{fig:specfifo-ifc}).
Now we explain each interface method:
\begin{itemize}
    \item Methods \code{enq}, \code{deq} and \code{first}: are the common FIFO interface methods.
    \item Method \code{first\_poisoned}: returns if the head of FIFO is poisoned or not.
    \item Subinterface \code{specUpdate}: See Section~\ref{sec:specupdate}.
\end{itemize}
If the module parameter \code{lazyEnq} is true, then method \code{deq} does not affect the guard of method \code{enq}, i.e., method \code{enq} cannot claim the dequeuing entry.

\begin{figure}[t]
\begin{lstlisting}[caption={}]
interface SpecPoisonFifo#(numeric type size, type t);
  method Action enq(ToSpecFifo#(t) x);
  method Action deq;
  method ToSpecFifo#(t) first_data;
  method Bool first_poisoned;
  interface SpeculationUpdate specUpdate;
endinterface
module mkSpecPoisonFifo#(Bool lazyEnq)(SpecPoisonFifo#(size, t));
  // module implementation
endmodule
\end{lstlisting}
\caption{Interface of speculative poisoned FIFO}\label{fig:specpoisonfifo-ifc}
\end{figure}

\noindent\textbf{Conflict Matrix:}
The conflict matrix of the interface methods is:
\begin{itemize}
    \item \code{deq} $<$ \code{enq} $<$ \code{correctSpeculation}
    \item \code{incorrectSpeculation} $<$ \code{correctSpeculation}
    \item \code{incorrectSpeculation} C \code{enq}
    \item \code{deq} $<$ \code{incorrectSpeculation}
\end{itemize}

\subsubsection{Implementation}
We use EHRs to store the fields for each FIFO entry.
Methods access EHRs using the appropriate ports according to the conflict matrix.

If the module parameter \code{lazyEnq} (Figure~\ref{fig:specfifo-ifc}) is true, then the guard of method \code{enq} is computed using the wire which is set using port 0 of all EHRs.
This cuts off the path from method \code{deq} to \code{enq}.

\subsubsection{Source Code}
See module \code{mkSpecPoisonFifo} in \code{//procs/procs/lib/SpecPoisonFifo.bsv}.
